---
title: "Social Media Impact on Teen Mental Health — Deep Analysis"
author: "Social Media Data Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
    theme: cosmo
    highlight: tango
    df_print: paged
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 7,
  fig.align  = "center",
  dpi     = 150
)
```

---

> **Dataset:** `teen_mental_health.csv` — 13 columns covering social media habits,
> sleep, academics, physical activity, and psychosocial outcomes in teenagers.
> **Target variable:** `depression_label` (binary: No Depression / Depression).

---

# Section 0 — Setup {.tabset}

## Install & Load Packages

```{r libraries}
# ── Install any missing packages automatically ──────────────────────────────
required_pkgs <- c(
  "tidyverse", "scales", "corrplot", "patchwork", "plotly",
  "caret", "randomForest", "xgboost", "e1071", "glmnet",
  "pROC", "vip", "GGally", "ggridges", "viridis",
  "kableExtra", "DT", "gridExtra"
)

installed <- rownames(installed.packages())
to_install <- setdiff(required_pkgs, installed)
if (length(to_install) > 0) {
  install.packages(to_install, repos = "https://cloud.r-project.org", quiet = TRUE)
}

# ── Load libraries ───────────────────────────────────────────────────────────
suppressPackageStartupMessages({
  library(tidyverse)
  library(scales)
  library(corrplot)
  library(patchwork)
  library(plotly)
  library(caret)
  library(randomForest)
  library(xgboost)
  library(e1071)
  library(glmnet)
  library(pROC)
  library(vip)
  library(GGally)
  library(ggridges)
  library(viridis)
  library(kableExtra)
  library(DT)
  library(gridExtra)
})

set.seed(42)

# ── Global theme ─────────────────────────────────────────────────────────────
theme_set(
  theme_minimal(base_size = 13) +
    theme(
      plot.title    = element_text(face = "bold", size = 14, hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
      axis.title    = element_text(size = 11),
      legend.position = "bottom",
      strip.text    = element_text(face = "bold")
    )
)

# ── Column definitions ────────────────────────────────────────────────────────
ALL_COLS <- c(
  "age", "gender", "daily_social_media_hours", "platform_usage",
  "sleep_hours", "screen_time_before_sleep", "academic_performance",
  "physical_activity", "social_interaction_level", "stress_level",
  "anxiety_level", "addiction_level", "depression_label"
)

NUMERIC_COLS <- c(
  "age", "daily_social_media_hours", "sleep_hours",
  "screen_time_before_sleep", "academic_performance",
  "physical_activity", "social_interaction_level",
  "stress_level", "anxiety_level", "addiction_level"
)

CAT_COLS    <- c("gender", "platform_usage")
TARGET_COL  <- "depression_label"

cat("✓ All packages loaded.\n")
cat("✓ Columns defined:", length(ALL_COLS), "total.\n")
```

## Load Data

```{r load-data}
# ── Read CSV (adjust path as needed) ─────────────────────────────────────────
# Place teen_mental_health.csv in a 'data/' subfolder OR change the path below.
csv_path <- "data/teen_mental_health.csv"

if (!file.exists(csv_path)) {
  stop(
    "❌  File not found: ", csv_path,
    "\n   Place your CSV in a 'data/' subfolder next to this .Rmd file."
  )
}

df_raw <- read_csv(csv_path, show_col_types = FALSE)

# Keep only the 13 specified columns (robust to extra columns in the CSV)
missing_cols <- setdiff(ALL_COLS, colnames(df_raw))
if (length(missing_cols) > 0) {
  stop("❌  These expected columns are missing from the CSV: ",
       paste(missing_cols, collapse = ", "))
}

df <- df_raw |> dplyr::select(all_of(ALL_COLS))

cat("✅  Dataset loaded successfully.\n")
cat("Shape:", nrow(df), "rows ×", ncol(df), "columns\n\n")

# Data types overview
glimpse(df)
```

```{r summary-table}
cat("\n── Descriptive Statistics ──────────────────────────────────────────────\n")
summary(df)
```

```{r head-table}
df |>
  head(10) |>
  kable(caption = "First 10 Rows of the Dataset") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 12)
```

---

# Section 1 — Data Cleaning & Preprocessing

## Missing Values

```{r missing-values, fig.height=5}
missing_counts <- colSums(is.na(df))
missing_df <- tibble(
  Column  = names(missing_counts),
  Missing = as.integer(missing_counts),
  Pct     = round(100 * missing_counts / nrow(df), 2)
) |> arrange(desc(Missing))

missing_df |>
  kable(caption = "Missing Values per Column") |>
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE, font_size = 12)
```

```{r missing-heatmap, fig.height=5, fig.width=12}
miss_long <- df |>
  mutate(row_id = row_number()) |>
  mutate(across(all_of(ALL_COLS), ~ as.integer(is.na(.)))) |>
  pivot_longer(-row_id, names_to = "column", values_to = "is_missing") |>
  mutate(column = factor(column, levels = ALL_COLS))

ggplot(miss_long, aes(column, row_id, fill = factor(is_missing))) +
  geom_tile() +
  scale_fill_manual(
    values = c("0" = "grey92", "1" = "firebrick"),
    labels = c("Present", "Missing"), name = NULL
  ) +
  scale_y_reverse() +
  labs(title = "Missing Values Heatmap", x = NULL, y = "Row Index") +
  theme(
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.text.y  = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid   = element_blank()
  )
```

> **Note:** If there are missing values, the code below imputes numeric columns
> with their median and categorical columns with their mode.

```{r impute-missing}
# Impute numeric columns with median
for (col in NUMERIC_COLS) {
  if (any(is.na(df[[col]]))) {
    df[[col]][is.na(df[[col]])] <- median(df[[col]], na.rm = TRUE)
    cat("Imputed median for:", col, "\n")
  }
}

# Impute categorical columns with mode
mode_val <- function(x) names(sort(table(x), decreasing = TRUE))[1]
for (col in c(CAT_COLS, TARGET_COL)) {
  if (any(is.na(df[[col]]))) {
    df[[col]][is.na(df[[col]])] <- mode_val(df[[col]])
    cat("Imputed mode for:", col, "\n")
  }
}

cat("\n✅  No missing values remain:", sum(is.na(df)), "\n")
```

## Remove Duplicates

```{r remove-duplicates}
n_before <- nrow(df)
df <- df |> distinct()
n_removed <- n_before - nrow(df)
cat("Rows before deduplication:", n_before, "\n")
cat("Duplicate rows removed   :", n_removed, "\n")
cat("Rows after deduplication :", nrow(df), "\n")
```

## Encode Target & Categoricals

```{r encode-target}
# ── Normalize depression_label to a clean binary factor ──────────────────────
unique_labels <- unique(as.character(df$depression_label))
cat("Unique depression_label values:", paste(sort(unique_labels), collapse = ", "), "\n")

# Handle both numeric (0/1) and string variants
df <- df |>
  mutate(
    depression_label = case_when(
      tolower(as.character(depression_label)) %in%
        c("1", "yes", "depression", "depressed", "true") ~ "Depression",
      tolower(as.character(depression_label)) %in%
        c("0", "no", "no depression", "not depressed", "false") ~ "No Depression",
      TRUE ~ as.character(depression_label)
    ),
    depression_label = factor(depression_label,
                              levels = c("No Depression", "Depression"))
  )

cat("Depression label levels:", levels(df$depression_label), "\n")

# ── Encode categoricals for numeric analysis ─────────────────────────────────
df <- df |>
  mutate(
    depression_label_enc = as.integer(depression_label == "Depression"),
    gender_enc           = as.integer(factor(gender)) - 1L,
    platform_usage_enc   = as.integer(factor(platform_usage)) - 1L,
    # social_interaction_level may be text (Low/Medium/High) or numeric
    social_int_enc = if (is.character(social_interaction_level) ||
                         is.factor(social_interaction_level)) {
      as.integer(factor(social_interaction_level,
                        levels = c("Low", "Medium", "High"))) - 1L
    } else {
      as.integer(social_interaction_level)
    }
  )
```

## Class Distribution

```{r class-dist}
cat("\n── Value Counts ─────────────────────────────────────────────────────────\n")
cat("\ngender:\n"); print(table(df$gender))
cat("\nplatform_usage:\n"); print(table(df$platform_usage))
cat("\ndepression_label:\n"); print(table(df$depression_label))
cat("\nClass balance (%):\n")
print(round(prop.table(table(df$depression_label)) * 100, 2))
```

---

# Section 2 — Exploratory Data Analysis (EDA)

## Distributions of Numeric Features

```{r dist-grid, fig.width=20, fig.height=9}
plots_dist <- lapply(NUMERIC_COLS, function(col) {
  vals <- if (col == "social_interaction_level") df$social_int_enc else df[[col]]
  tmp  <- tibble(x = vals)

  ggplot(tmp, aes(x)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30,
                   fill = "#4E79A7", colour = "white", alpha = 0.8) +
    geom_density(colour = "#E15759", linewidth = 1) +
    labs(title = col, x = NULL, y = "Density")
})

wrap_plots(plots_dist, ncol = 5) +
  plot_annotation(
    title    = "Distributions of All Numeric Features",
    subtitle = "Blue = histogram  |  Red = KDE curve",
    theme    = theme(plot.title = element_text(face = "bold", size = 16, hjust = 0.5))
  )
```

## Countplots — Categorical Features

```{r countplots, fig.width=16, fig.height=5}
p_gender <- ggplot(df, aes(gender, fill = gender)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.8) +
  labs(title = "Gender Distribution", x = NULL, y = "Count")

p_platform <- ggplot(df, aes(platform_usage, fill = platform_usage)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  scale_fill_viridis_d(option = "C", begin = 0.1, end = 0.9) +
  labs(title = "Platform Usage Distribution", x = NULL, y = "Count")

p_dep <- ggplot(df, aes(depression_label, fill = depression_label)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Depression Label Distribution", x = NULL, y = "Count")

p_gender + p_platform + p_dep +
  plot_annotation(title = "Categorical Feature Distributions",
                  theme = theme(plot.title = element_text(face = "bold", size = 15, hjust = 0.5)))
```

## Pearson Correlation Heatmap

```{r corr-heatmap, fig.width=12, fig.height=10}
corr_data <- df |>
  transmute(
    age, daily_social_media_hours, sleep_hours,
    screen_time_before_sleep, academic_performance,
    physical_activity, social_interaction = social_int_enc,
    stress_level, anxiety_level, addiction_level,
    depression = depression_label_enc
  ) |>
  # Force all columns to numeric (handles any hidden factor/character columns)
  mutate(across(everything(), ~ suppressWarnings(as.numeric(as.character(.))))) |>
  # Drop any column that is entirely NA after coercion
  select(where(~ sum(!is.na(.)) > 1))

# Verify we have usable data
if (ncol(corr_data) < 2) {
  stop("Not enough numeric columns for correlation matrix.")
}

# Use pairwise.complete.obs as a safer fallback
corr_mat <- tryCatch(
  cor(corr_data, use = "complete.obs",  method = "pearson"),
  error = function(e) cor(corr_data, use = "pairwise.complete.obs", method = "pearson")
)

# Sync row/col names for corrplot
rownames(corr_mat) <- colnames(corr_mat)

corr_mat <- cor(corr_data, use = "complete.obs", method = "pearson")

corrplot(
  corr_mat,
  method      = "color",
  type        = "upper",
  order       = "hclust",
  addCoef.col = "black",
  number.cex  = 0.65,
  tl.cex      = 0.85,
  tl.col      = "black",
  col         = colorRampPalette(c("#2166AC", "white", "#D6604D"))(200),
  title       = "Pearson Correlation Heatmap (All Numeric Features)",
  mar         = c(0, 0, 2, 0)
)
```

> **Reading the heatmap:** Red = positive correlation, Blue = negative.
> Darker shades = stronger relationship. Look for strong correlates of `depression`.

## Pairplot — Key Mental Health Variables

```{r pairplot, fig.width=14, fig.height=12}
pair_cols <- c(
  "daily_social_media_hours", "sleep_hours",
  "stress_level", "anxiety_level", "addiction_level"
)

ggpairs(
  df[, c(pair_cols, "depression_label")],
  aes(colour = depression_label, alpha = 0.4),
  upper = list(continuous = wrap("cor", size = 3.5)),
  lower = list(continuous = wrap("points", alpha = 0.3, size = 0.8)),
  diag  = list(continuous = wrap("densityDiag", alpha = 0.5)),
  columnLabels = c("Social Media (h)", "Sleep (h)",
                   "Stress", "Anxiety", "Addiction", "Depression")
) +
  scale_colour_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  scale_fill_manual(values  = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Pairplot — Key Mental Health Variables by Depression Label") +
  theme(axis.text = element_text(size = 7))
```

## Boxplots by Depression Label

```{r box-depression, fig.width=18, fig.height=10}
box_cols   <- c("stress_level", "anxiety_level", "addiction_level",
                "sleep_hours", "daily_social_media_hours",
                "screen_time_before_sleep", "academic_performance",
                "physical_activity")

box_labels <- c("Stress Level", "Anxiety Level", "Addiction Level",
                "Sleep Hours", "Social Media (h/day)",
                "Screen Time Before Sleep", "Academic Performance",
                "Physical Activity")

bp_list <- Map(function(col, lbl) {
  ggplot(df, aes(depression_label, .data[[col]], fill = depression_label)) +
    geom_boxplot(alpha = 0.75, outlier.shape = 21, outlier.size = 1.5) +
    geom_jitter(width = 0.1, alpha = 0.08, size = 0.6) +
    scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759"),
                      guide  = "none") +
    labs(title = lbl, x = NULL, y = lbl)
}, box_cols, box_labels)

wrap_plots(bp_list, ncol = 4) +
  plot_annotation(
    title    = "Feature Distributions by Depression Label",
    subtitle = "Green = No Depression  |  Red = Depression",
    theme    = theme(plot.title = element_text(face = "bold", size = 15, hjust = 0.5))
  )
```

## Ridge Plots — Overlapping Distributions

```{r ridge-plots, fig.width=14, fig.height=9}
ridge_cols <- c("stress_level", "anxiety_level", "addiction_level",
                "sleep_hours", "daily_social_media_hours")

df_ridge <- df |>
  select(all_of(ridge_cols), depression_label) |>
  pivot_longer(-depression_label, names_to = "feature", values_to = "value")

ggplot(df_ridge, aes(value, feature, fill = depression_label, colour = depression_label)) +
  geom_density_ridges(alpha = 0.55, scale = 1.2, rel_min_height = 0.01) +
  scale_fill_manual(values   = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  scale_colour_manual(values = c("No Depression" = "#2D6A2E", "Depression" = "#A01B1B")) +
  labs(
    title    = "Ridge Density Plots by Depression Label",
    subtitle = "Overlap reveals how well each feature separates the two classes",
    x = "Value", y = NULL, fill = "Label", colour = "Label"
  )
```

---

# Section 3 — Deep Statistical Analysis

## IQR Outlier Detection

```{r iqr-outliers, fig.width=20, fig.height=8}
iqr_count <- function(x) {
  q  <- quantile(x, c(0.25, 0.75), na.rm = TRUE)
  iq <- q[2] - q[1]
  sum(x < q[1] - 1.5 * iq | x > q[2] + 1.5 * iq, na.rm = TRUE)
}

outlier_df <- tibble(
  Feature  = NUMERIC_COLS,
  Outliers = sapply(NUMERIC_COLS, function(c) {
    vals <- if (c == "social_interaction_level") df$social_int_enc else df[[c]]
    iqr_count(vals)
  }),
  Pct = round(Outliers / nrow(df) * 100, 2)
) |> arrange(desc(Outliers))

outlier_df |>
  kable(caption = "Outlier Counts by IQR Method") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r outlier-boxplots, fig.width=20, fig.height=8}
obox_list <- lapply(NUMERIC_COLS, function(col) {
  vals <- if (col == "social_interaction_level") df$social_int_enc else df[[col]]
  tmp  <- tibble(x = vals)
  ggplot(tmp, aes(y = x)) +
    geom_boxplot(fill = "#4E79A7", alpha = 0.7, outlier.colour = "red", outlier.size = 1.5) +
    labs(title = col, x = NULL, y = col)
})

wrap_plots(obox_list, ncol = 5) +
  plot_annotation(
    title = "IQR Outlier Detection — All Numeric Features",
    theme = theme(plot.title = element_text(face = "bold", size = 15, hjust = 0.5))
  )
```

## T-Tests: Depressed vs Non-Depressed

```{r ttest-helper}
# Helper: run Welch t-test and return a tidy tibble row
run_ttest <- function(col_name, df) {
  g0 <- df |> filter(depression_label == "No Depression") |> pull(!!sym(col_name))
  g1 <- df |> filter(depression_label == "Depression")    |> pull(!!sym(col_name))
  tt <- t.test(g0, g1, var.equal = FALSE)
  tibble(
    Feature     = col_name,
    Mean_NoDep  = round(mean(g0, na.rm = TRUE), 3),
    Mean_Dep    = round(mean(g1, na.rm = TRUE), 3),
    t_statistic = round(tt$statistic, 4),
    p_value     = signif(tt$p.value, 4),
    Significant = ifelse(tt$p.value < 0.05, "✅ Yes", "❌ No")
  )
}

ttest_cols <- c("daily_social_media_hours", "sleep_hours",
                "anxiety_level", "addiction_level",
                "stress_level", "screen_time_before_sleep",
                "academic_performance", "physical_activity")

ttest_results <- bind_rows(lapply(ttest_cols, run_ttest, df = df))

ttest_results |>
  kable(caption = "Welch T-Tests: No Depression vs Depression Groups") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 12) |>
  column_spec(6, bold = TRUE,
              color = ifelse(ttest_results$Significant == "✅ Yes", "green", "red"))
```

> **Interpretation:** Features marked ✅ have statistically significant mean differences
> between depressed and non-depressed teens (p < 0.05). Compare `Mean_NoDep` vs
> `Mean_Dep` columns to see the direction of the effect.

```{r ttest-viz, fig.width=14, fig.height=5}
ttest_plot_df <- ttest_results |>
  mutate(neg_log_p = -log10(p_value)) |>
  arrange(desc(neg_log_p))

ggplot(ttest_plot_df, aes(reorder(Feature, neg_log_p), neg_log_p,
                           fill = Significant)) +
  geom_col(alpha = 0.85) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed",
             colour = "red", linewidth = 0.8) +
  scale_fill_manual(values = c("✅ Yes" = "#59A14F", "❌ No" = "#BAB0AC")) +
  coord_flip() +
  labs(
    title    = "T-Test Significance (-log10 p-value)",
    subtitle = "Red dashed line = p = 0.05 threshold",
    x = NULL, y = "-log10(p-value)", fill = "Significant"
  )
```

## ANOVA — Platform Usage Effect

```{r anova-tests}
run_anova <- function(outcome, df) {
  formula <- as.formula(paste(outcome, "~ platform_usage"))
  aov_fit <- aov(formula, data = df)
  sm      <- summary(aov_fit)[[1]]
  tibble(
    Outcome     = outcome,
    F_statistic = round(sm[1, "F value"], 4),
    p_value     = signif(sm[1, "Pr(>F)"], 4),
    Significant = ifelse(sm[1, "Pr(>F)"] < 0.05, "✅ Yes", "❌ No")
  )
}

anova_outcomes <- c("stress_level", "anxiety_level", "addiction_level",
                    "sleep_hours", "daily_social_media_hours")

anova_results <- bind_rows(lapply(anova_outcomes, run_anova, df = df))

anova_results |>
  kable(caption = "One-Way ANOVA: Outcomes by Platform Usage") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

> **Interpretation:** A significant ANOVA (p < 0.05) means mean scores differ across
> at least two platform groups (Instagram, TikTok, Both). Post-hoc Tukey HSD would
> identify which pairs differ.

```{r tukey-hsd}
# Tukey HSD post-hoc for stress_level (most relevant outcome)
aov_stress_full <- aov(stress_level ~ platform_usage, data = df)
cat("Tukey HSD post-hoc for stress_level ~ platform_usage:\n")
print(TukeyHSD(aov_stress_full))
```

## Chi-Square — Gender vs Depression

```{r chi-square}
ct <- table(df$gender, df$depression_label)
cat("Contingency table:\n"); print(ct)

chi_test <- chisq.test(ct)
cat("\nChi-square test:\n"); print(chi_test)

chi_interp <- if (chi_test$p.value < 0.05) {
  "✅  Significant: gender and depression_label are NOT independent (p < 0.05)."
} else {
  "❌  Not significant: no strong association between gender and depression_label."
}
cat("\n", chi_interp, "\n")
```

```{r chi-viz, fig.width=8, fig.height=5}
as.data.frame(ct) |>
  rename(Gender = Var1, Depression = Var2, Count = Freq) |>
  group_by(Gender) |>
  mutate(Prop = Count / sum(Count)) |>
  ggplot(aes(Gender, Prop, fill = Depression)) +
  geom_col(position = "fill", alpha = 0.85) +
  geom_text(aes(label = scales::percent(Prop, accuracy = 0.1)),
            position = position_fill(vjust = 0.5), size = 4) +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Depression Label Proportion by Gender",
       x = NULL, y = "Proportion", fill = NULL)
```

## Spearman Correlations with Depression

```{r spearman}
spearman_cols <- c(
  "daily_social_media_hours", "sleep_hours", "screen_time_before_sleep",
  "stress_level", "anxiety_level", "addiction_level",
  "academic_performance", "physical_activity", "age"
)

spearman_results <- map_dfr(spearman_cols, function(col) {
  test <- cor.test(df$depression_label_enc, df[[col]],
                   method = "spearman", exact = FALSE)
  tibble(
    Feature     = col,
    Spearman_rho = round(unname(test$estimate), 4),
    p_value      = signif(test$p.value, 4),
    Significant  = ifelse(test$p.value < 0.05, "✅ Yes", "❌ No"),
    Direction    = ifelse(unname(test$estimate) > 0, "Positive ↑", "Negative ↓")
  )
}) |> arrange(desc(abs(Spearman_rho)))

spearman_results |>
  kable(caption = "Spearman Correlations: Feature vs Depression Label (Encoded)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  column_spec(2, bold = TRUE,
              color = ifelse(spearman_results$Spearman_rho > 0, "#E15759", "#4E79A7"))
```

```{r spearman-viz, fig.width=10, fig.height=5}
ggplot(spearman_results,
       aes(reorder(Feature, Spearman_rho), Spearman_rho,
           fill = Spearman_rho > 0)) +
  geom_col(alpha = 0.85) +
  geom_hline(yintercept = 0, colour = "black", linewidth = 0.5) +
  scale_fill_manual(values = c("TRUE" = "#E15759", "FALSE" = "#4E79A7"),
                    labels = c("Negative", "Positive"), name = "Direction") +
  coord_flip() +
  labs(
    title    = "Spearman Correlations with Depression Label",
    subtitle = "Red = positive correlation  |  Blue = negative",
    x = NULL, y = "Spearman ρ"
  )
```

---

# Section 4 — Feature Engineering

```{r feature-engineering}
df <- df |>
  mutate(
    # --- Derived numeric features ---
    sleep_deficit       = 8 - sleep_hours,
    screen_load         = daily_social_media_hours + screen_time_before_sleep,

    # --- Composite mental health score (MinMax normalised sum) ---
    stress_n            = (stress_level  - min(stress_level))  / (max(stress_level)  - min(stress_level)),
    anxiety_n           = (anxiety_level - min(anxiety_level)) / (max(anxiety_level) - min(anxiety_level)),
    addiction_n         = (addiction_level - min(addiction_level)) / (max(addiction_level) - min(addiction_level)),
    mental_health_score = stress_n + anxiety_n + addiction_n,

    # --- Usage category bins ---
    usage_category = cut(
      daily_social_media_hours,
      breaks = c(-Inf, 2, 4, Inf),
      labels = c("Light (<2h)", "Moderate (2-4h)", "Heavy (>4h)"),
      right  = FALSE
    ),

    # --- Age group bins ---
    age_group = cut(
      age,
      breaks         = c(12, 15, 17, 19),
      labels         = c("Early Teen (13-15)", "Mid Teen (16-17)", "Late Teen (18-19)"),
      include.lowest = TRUE
    )
  ) |>
  # Drop helper normalised columns
  select(-stress_n, -anxiety_n, -addiction_n)

cat("✅  Feature engineering complete.\n")
cat("New columns added: sleep_deficit, screen_load, mental_health_score,",
    "usage_category, age_group\n\n")

df |>
  select(sleep_deficit, screen_load, mental_health_score, usage_category, age_group) |>
  head(8) |>
  kable(caption = "Sample of Engineered Features") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r usage-age-dist, fig.width=14, fig.height=5}
p_uc <- ggplot(df, aes(usage_category, fill = usage_category)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  scale_fill_viridis_d(option = "B", begin = 0.2, end = 0.85) +
  labs(title = "Usage Category Distribution", x = NULL, y = "Count")

p_ag <- ggplot(df, aes(age_group, fill = age_group)) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.4) +
  scale_fill_viridis_d(option = "E", begin = 0.15, end = 0.9) +
  labs(title = "Age Group Distribution", x = NULL, y = "Count")

p_uc + p_ag
```

## Build Model Matrix

```{r model-matrix}
# Encode categoricals for modeling
df_enc <- df |>
  mutate(
    gender_enc    = as.integer(factor(gender)) - 1L,
    platform_enc  = as.integer(factor(platform_usage)) - 1L,
    usage_cat_enc = as.integer(usage_category) - 1L,
    age_group_enc = as.integer(age_group) - 1L
  )

feature_cols <- c(
  "age", "daily_social_media_hours", "sleep_hours", "screen_time_before_sleep",
  "academic_performance", "physical_activity", "social_int_enc",
  "stress_level", "anxiety_level", "addiction_level",
  "sleep_deficit", "screen_load", "mental_health_score",
  "gender_enc", "platform_enc", "usage_cat_enc", "age_group_enc"
)

# Force all feature columns to numeric
X_raw <- df_enc |>
  select(all_of(feature_cols)) |>
  mutate(across(everything(), ~ suppressWarnings(as.numeric(as.character(.))))) |>
  as.matrix()

# Replace any remaining NA / NaN / Inf with column medians
for (j in seq_len(ncol(X_raw))) {
  bad <- !is.finite(X_raw[, j])
  if (any(bad)) {
    med <- median(X_raw[!bad, j], na.rm = TRUE)
    if (!is.finite(med)) med <- 0
    X_raw[bad, j] <- med
  }
}

X_scaled <- scale(X_raw)
# scale() returns NaN for zero-variance columns — glmnet rejects those
X_scaled[!is.finite(X_scaled)] <- 0
colnames(X_scaled) <- feature_cols

y_bin <- df$depression_label_enc

cat("✅  Feature matrix shape:", nrow(X_scaled), "rows ×", ncol(X_scaled), "columns\n")
cat("Any non-finite values remaining:", any(!is.finite(X_scaled)), "\n")
cat("Target distribution (0 = No Depression, 1 = Depression):\n")
print(table(y_bin))
```

---

# Section 5 — Advanced Visualisations

## Interactive: Social Media Hours vs Anxiety

```{r plotly-scatter, fig.width=10, fig.height=6}
p_interactive <- df |>
  plot_ly(
    x    = ~daily_social_media_hours,
    y    = ~anxiety_level,
    color = ~depression_label,
    size  = ~addiction_level,
    text  = ~paste0(
      "Platform: ", platform_usage,
      "<br>Sleep: ", sleep_hours, "h",
      "<br>Stress: ", stress_level,
      "<br>Gender: ", gender
    ),
    type   = "scatter",
    mode   = "markers",
    colors = c("No Depression" = "#59A14F", "Depression" = "#E15759"),
    marker = list(opacity = 0.65, sizemode = "diameter", sizeref = 0.25)
  ) |>
  layout(
    title = list(text = "Social Media Hours vs Anxiety Level",
                 font = list(size = 16)),
    xaxis = list(title = "Daily Social Media Hours"),
    yaxis = list(title = "Anxiety Level"),
    legend = list(title = list(text = "Depression Label"))
  )
p_interactive
```

## Interactive: Stress & Anxiety by Platform

```{r plotly-platform, fig.width=10, fig.height=6}
df_long_plat <- df |>
  select(platform_usage, stress_level, anxiety_level, addiction_level) |>
  pivot_longer(-platform_usage, names_to = "metric", values_to = "score")

plot_ly(df_long_plat,
        x = ~platform_usage, y = ~score, color = ~metric,
        type = "box", boxpoints = "outliers") |>
  layout(
    title  = list(text = "Mental Health Scores by Platform Usage"),
    xaxis  = list(title = "Platform"),
    yaxis  = list(title = "Score"),
    boxmode = "group"
  )
```

## Violin Plots

```{r violin-plots, fig.width=14, fig.height=5}
vp1 <- ggplot(df, aes(depression_label, sleep_hours, fill = depression_label)) +
  geom_violin(trim = FALSE, alpha = 0.65) +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759"),
                    guide = "none") +
  labs(title = "Sleep Hours by Depression Label", x = NULL, y = "Sleep Hours")

vp2 <- ggplot(df, aes(depression_label, addiction_level, fill = depression_label)) +
  geom_violin(trim = FALSE, alpha = 0.65) +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759"),
                    guide = "none") +
  labs(title = "Addiction Level by Depression Label", x = NULL, y = "Addiction Level")

vp3 <- ggplot(df, aes(depression_label, stress_level, fill = depression_label)) +
  geom_violin(trim = FALSE, alpha = 0.65) +
  geom_boxplot(width = 0.08, fill = "white", outlier.shape = NA) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759"),
                    guide = "none") +
  labs(title = "Stress Level by Depression Label", x = NULL, y = "Stress Level")

vp1 + vp2 + vp3 +
  plot_annotation(title = "Violin Plots: Key Features by Depression Label",
                  theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
```

## Heatmaps — Pivot Tables

```{r heatmap-anxiety, fig.width=10, fig.height=5}
pivot_anx <- df |>
  filter(!is.na(age_group), !is.na(usage_category)) |>
  group_by(age_group, usage_category) |>
  summarise(mean_anxiety = round(mean(anxiety_level), 2), .groups = "drop")

ggplot(pivot_anx, aes(usage_category, age_group, fill = mean_anxiety)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = mean_anxiety), colour = "black", size = 4.5) +
  scale_fill_gradient(low = "#FFF5EB", high = "#D94801",
                      name = "Mean\nAnxiety") +
  labs(title = "Mean Anxiety Level: Age Group × Usage Category",
       x = "Usage Category", y = "Age Group")
```

```{r heatmap-stress, fig.width=10, fig.height=5}
pivot_stress <- df |>
  group_by(gender, platform_usage) |>
  summarise(mean_stress = round(mean(stress_level), 2), .groups = "drop")

ggplot(pivot_stress, aes(platform_usage, gender, fill = mean_stress)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = mean_stress), colour = "black", size = 4.5) +
  scale_fill_gradient(low = "#EFF3FF", high = "#2171B5",
                      name = "Mean\nStress") +
  labs(title = "Mean Stress Level: Gender × Platform Usage",
       x = "Platform Usage", y = "Gender")
```

## Stacked Bar Charts

```{r stacked-bars, fig.width=14, fig.height=5}
sb1 <- df |>
  count(gender, depression_label) |>
  group_by(gender) |>
  mutate(prop = n / sum(n)) |>
  ggplot(aes(gender, prop, fill = depression_label)) +
  geom_col(position = "fill", alpha = 0.85) +
  geom_text(aes(label = percent(prop, accuracy = 0.1)),
            position = position_fill(vjust = 0.5), size = 3.5) +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Depression by Gender", x = NULL, y = "Proportion", fill = NULL)

sb2 <- df |>
  filter(!is.na(usage_category)) |>
  count(usage_category, depression_label) |>
  group_by(usage_category) |>
  mutate(prop = n / sum(n)) |>
  ggplot(aes(usage_category, prop, fill = depression_label)) +
  geom_col(position = "fill", alpha = 0.85) +
  geom_text(aes(label = percent(prop, accuracy = 0.1)),
            position = position_fill(vjust = 0.5), size = 3.5) +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Depression by Usage Category", x = NULL, y = "Proportion", fill = NULL)

sb1 + sb2 +
  plot_annotation(title = "Depression Label Distribution",
                  theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
```

## Line Plot — Mental Health Trend Across Usage

```{r line-usage, fig.width=10, fig.height=5}
line_df <- df |>
  filter(!is.na(usage_category)) |>
  group_by(usage_category) |>
  summarise(
    `Stress Level`    = mean(stress_level),
    `Anxiety Level`   = mean(anxiety_level),
    `Addiction Level` = mean(addiction_level),
    .groups = "drop"
  ) |>
  pivot_longer(-usage_category, names_to = "Metric", values_to = "Mean Score")

ggplot(line_df, aes(usage_category, `Mean Score`, colour = Metric, group = Metric)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  scale_colour_manual(values = c(
    "Stress Level"    = "#E15759",
    "Anxiety Level"   = "#F28E2B",
    "Addiction Level" = "#4E79A7"
  )) +
  labs(
    title    = "Mean Mental Health Scores Across Usage Categories",
    subtitle = "Heavier social media use → higher stress, anxiety, and addiction",
    x = "Usage Category", y = "Mean Score", colour = "Metric"
  )
```

## Regression Scatter Plots

```{r regression-scatters, fig.width=14, fig.height=5}
rs1 <- ggplot(df, aes(screen_time_before_sleep, sleep_hours,
                       colour = depression_label)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_colour_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Screen Time Before Sleep vs Sleep Hours",
       x = "Screen Time Before Sleep (h)", y = "Sleep Hours", colour = NULL)

rs2 <- ggplot(df, aes(academic_performance, mental_health_score,
                       colour = depression_label)) +
  geom_point(alpha = 0.3, size = 1.2) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_colour_manual(values = c("No Depression" = "#59A14F", "Depression" = "#E15759")) +
  labs(title = "Academic Performance vs Mental Health Score",
       x = "Academic Performance", y = "Mental Health Score", colour = NULL)

rs1 + rs2 +
  plot_annotation(title = "Regression Relationships by Depression Label",
                  theme = theme(plot.title = element_text(face = "bold", hjust = 0.5)))
```

---

# Section 6 — Machine Learning

## Train / Test Split

```{r ml-split}
# Ensure train/test matrices are fully finite (required by glmnet)
clean_matrix <- function(m) {
  m <- as.matrix(m)
  m[!is.finite(m)] <- 0
  m
}

y_fac       <- factor(y_bin, levels = c(0, 1),
                      labels = c("NoDepression", "Depression"))
train_idx   <- createDataPartition(y_fac, p = 0.8, list = FALSE)
X_train     <- clean_matrix(X_scaled[train_idx, , drop = FALSE])
X_test      <- clean_matrix(X_scaled[-train_idx, , drop = FALSE])
y_train_fac <- y_fac[train_idx]
y_test_fac  <- y_fac[-train_idx]

cat("Train size:", nrow(X_train), " | Test size:", nrow(X_test), "\n")
cat("Train class balance:\n"); print(table(y_train_fac))
cat("Test class balance:\n");  print(table(y_test_fac))
```

## Confusion Matrix Helper

```{r cm-helper}
plot_cm <- function(pred, actual, title) {
  cm  <- confusionMatrix(pred, actual, positive = "Depression")
  tab <- as.data.frame(cm$table)
  names(tab) <- c("Actual", "Predicted", "Freq")

  p <- ggplot(tab, aes(Predicted, Actual, fill = Freq)) +
    geom_tile(colour = "white") +
    geom_text(aes(label = Freq), size = 5, fontface = "bold") +
    scale_fill_gradient(low = "#EFF3FF", high = "#2171B5", name = "Count") +
    labs(title = title, x = "Predicted", y = "Actual") +
    theme(legend.position = "right")

  list(plot = p, cm = cm)
}
```

## Macro F1 Helper

```{r f1-helper}
# ── Fixed macro F1 (avoids MLmetrics 'average' argument issue) ───────────────
macro_f1 <- function(pred, actual) {
  lvls <- levels(actual)
  f1s  <- vapply(lvls, function(lv) {
    tp <- sum(pred == lv & actual == lv)
    fp <- sum(pred == lv & actual != lv)
    fn <- sum(pred != lv & actual == lv)
    if ((tp + fp) == 0 || (tp + fn) == 0) return(0)
    prec <- tp / (tp + fp)
    rec  <- tp / (tp + fn)
    if ((prec + rec) == 0) return(0)
    2 * prec * rec / (prec + rec)
  }, numeric(1))
  mean(f1s)
}
```

## Model Training & Evaluation

```{r train-lr}
cat(strrep("=", 60), "\n1. LOGISTIC REGRESSION (Ridge via glmnet)\n",
    strrep("=", 60), "\n")

stopifnot(
  "X_train has missing/non-finite values" =
    all(is.finite(X_train))
)

fit_lr <- cv.glmnet(
  X_train, as.integer(y_train_fac) - 1L,
  family = "binomial", alpha = 0, nfolds = 5
)
prob_lr  <- as.numeric(predict(fit_lr, X_test, type = "response", s = "lambda.min"))
pred_lr  <- factor(ifelse(prob_lr > 0.5, "Depression", "NoDepression"),
                   levels = levels(y_fac))

lr_cm    <- plot_cm(pred_lr, y_test_fac, "Logistic Regression — Confusion Matrix")
print(lr_cm$plot)
print(lr_cm$cm)

roc_lr   <- roc(as.integer(y_test_fac) - 1L, prob_lr, quiet = TRUE)
lr_auc   <- as.numeric(auc(roc_lr))
lr_acc   <- mean(pred_lr == y_test_fac)
lr_f1    <- macro_f1(pred_lr, y_test_fac)
cat(sprintf("Test Accuracy: %.4f | Macro F1: %.4f | ROC-AUC: %.4f\n",
            lr_acc, lr_f1, lr_auc))
```

```{r train-rf}
cat(strrep("=", 60), "\n2. RANDOM FOREST\n", strrep("=", 60), "\n")

fit_rf   <- randomForest(
  x = as.data.frame(X_train), y = y_train_fac,
  ntree = 200, importance = TRUE
)
pred_rf  <- predict(fit_rf, as.data.frame(X_test))
prob_rf  <- predict(fit_rf, as.data.frame(X_test), type = "prob")[, "Depression"]

rf_cm    <- plot_cm(pred_rf, y_test_fac, "Random Forest — Confusion Matrix")
print(rf_cm$plot)
print(rf_cm$cm)

roc_rf   <- roc(as.integer(y_test_fac) - 1L, prob_rf, quiet = TRUE)
rf_auc   <- as.numeric(auc(roc_rf))
rf_acc   <- mean(pred_rf == y_test_fac)
rf_f1    <- macro_f1(pred_rf, y_test_fac)
cat(sprintf("Test Accuracy: %.4f | Macro F1: %.4f | ROC-AUC: %.4f\n",
            rf_acc, rf_f1, rf_auc))
```

```{r train-xgb}
cat(strrep("=", 60), "\n3. XGBOOST\n", strrep("=", 60), "\n")

dtrain <- xgb.DMatrix(data = X_train, label = as.integer(y_train_fac) - 1L)
dtest  <- xgb.DMatrix(data = X_test,  label = as.integer(y_test_fac)  - 1L)

fit_xgb <- xgb.train(
  params = list(
    objective    = "binary:logistic",
    eval_metric  = "auc",
    eta          = 0.1,
    max_depth    = 6,
    subsample    = 0.8,
    colsample_bytree = 0.8
  ),
  data     = dtrain,
  nrounds  = 150,
  watchlist = list(train = dtrain, test = dtest),
  verbose  = 0
)

prob_xgb  <- predict(fit_xgb, dtest)
pred_xgb  <- factor(ifelse(prob_xgb > 0.5, "Depression", "NoDepression"),
                    levels = levels(y_fac))

xgb_cm    <- plot_cm(pred_xgb, y_test_fac, "XGBoost — Confusion Matrix")
print(xgb_cm$plot)
print(xgb_cm$cm)

roc_xgb   <- roc(as.integer(y_test_fac) - 1L, prob_xgb, quiet = TRUE)
xgb_auc   <- as.numeric(auc(roc_xgb))
xgb_acc   <- mean(pred_xgb == y_test_fac)
xgb_f1    <- macro_f1(pred_xgb, y_test_fac)
cat(sprintf("Test Accuracy: %.4f | Macro F1: %.4f | ROC-AUC: %.4f\n",
            xgb_acc, xgb_f1, xgb_auc))
```

```{r train-svm}
cat(strrep("=", 60), "\n4. SUPPORT VECTOR MACHINE (RBF)\n", strrep("=", 60), "\n")

fit_svm   <- svm(
  x = X_train, y = y_train_fac,
  kernel = "radial", probability = TRUE, scale = FALSE
)
pred_svm  <- predict(fit_svm, X_test)
prob_svm  <- attr(predict(fit_svm, X_test, probability = TRUE),
                  "probabilities")[, "Depression"]

svm_cm    <- plot_cm(pred_svm, y_test_fac, "SVM (RBF) — Confusion Matrix")
print(svm_cm$plot)
print(svm_cm$cm)

roc_svm   <- roc(as.integer(y_test_fac) - 1L, prob_svm, quiet = TRUE)
svm_auc   <- as.numeric(auc(roc_svm))
svm_acc   <- mean(pred_svm == y_test_fac)
svm_f1    <- macro_f1(pred_svm, y_test_fac)
cat(sprintf("Test Accuracy: %.4f | Macro F1: %.4f | ROC-AUC: %.4f\n",
            svm_acc, svm_f1, svm_auc))
```

## ROC Curves — All Models

```{r roc-curves, fig.width=9, fig.height=7}
plot(roc_lr,  col = "#4E79A7", lwd = 2,
     main = "ROC Curves — All Models")
plot(roc_rf,  col = "#59A14F", lwd = 2, add = TRUE)
plot(roc_xgb, col = "#E15759", lwd = 2, add = TRUE)
plot(roc_svm, col = "#F28E2B", lwd = 2, add = TRUE)
abline(a = 0, b = 1, lty = 2, col = "grey50")
legend(
  "bottomright",
  legend = c(
    sprintf("Logistic Regression (AUC = %.3f)", lr_auc),
    sprintf("Random Forest       (AUC = %.3f)", rf_auc),
    sprintf("XGBoost             (AUC = %.3f)", xgb_auc),
    sprintf("SVM (RBF)           (AUC = %.3f)", svm_auc)
  ),
  col  = c("#4E79A7", "#59A14F", "#E15759", "#F28E2B"),
  lwd  = 2,
  cex  = 0.9,
  bty  = "n"
)
```

## 5-Fold Cross-Validation

```{r cv-accuracy}
folds <- createFolds(y_fac, k = 5, list = TRUE, returnTrain = TRUE)

cv_rf <- vapply(folds, function(tr) {
  te <- setdiff(seq_len(nrow(X_scaled)), tr)
  m  <- randomForest(x = as.data.frame(X_scaled[tr, ]),
                     y = y_fac[tr], ntree = 100)
  mean(predict(m, as.data.frame(X_scaled[te, ])) == y_fac[te])
}, numeric(1))

cv_xgb <- vapply(folds, function(tr) {
  te   <- setdiff(seq_len(nrow(X_scaled)), tr)
  dtr  <- xgb.DMatrix(X_scaled[tr, ], label = as.integer(y_fac[tr]) - 1L)
  dte  <- xgb.DMatrix(X_scaled[te, ], label = as.integer(y_fac[te]) - 1L)
  m    <- xgb.train(
    params  = list(objective = "binary:logistic", eval_metric = "auc",
                   eta = 0.1, max_depth = 6),
    data    = dtr, nrounds = 100, verbose = 0
  )
  p    <- ifelse(predict(m, dte) > 0.5, "Depression", "NoDepression")
  mean(p == as.character(y_fac[te]))
}, numeric(1))

cat(sprintf(
  "RF  5-Fold CV Accuracy: %.4f ± %.4f\n",
  mean(cv_rf), sd(cv_rf)
))
cat(sprintf(
  "XGB 5-Fold CV Accuracy: %.4f ± %.4f\n",
  mean(cv_xgb), sd(cv_xgb)
))
```

## Model Comparison Table

```{r model-comparison, fig.width=12, fig.height=6}
comparison <- tibble(
  Model         = c("Logistic Regression", "Random Forest", "XGBoost", "SVM (RBF)"),
  `CV Accuracy` = c(NA, round(mean(cv_rf), 4), round(mean(cv_xgb), 4), NA),
  `Test Accuracy` = round(c(lr_acc, rf_acc, xgb_acc, svm_acc), 4),
  `F1-Macro`    = round(c(lr_f1,  rf_f1,  xgb_f1,  svm_f1),  4),
  `ROC-AUC`     = round(c(lr_auc, rf_auc, xgb_auc, svm_auc), 4)
) |> arrange(desc(`F1-Macro`))

comparison |>
  kable(caption = "Model Performance Summary (sorted by Macro F1)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) |>
  row_spec(1, bold = TRUE, background = "#E8F5E9")
```

```{r comparison-plot, fig.width=13, fig.height=6}
comparison |>
  pivot_longer(-Model, names_to = "Metric", values_to = "Score") |>
  filter(!is.na(Score)) |>
  ggplot(aes(Model, Score, fill = Metric)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_text(aes(label = round(Score, 3)),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3.2) +
  scale_fill_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  scale_y_continuous(limits = c(0, 1.1), labels = number_format(accuracy = 0.01)) +
  labs(
    title    = "Model Performance Comparison",
    subtitle = "Higher is better for all metrics",
    x = NULL, y = "Score", fill = "Metric"
  ) +
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
```

---

# Section 7 — Model Explainability

```{r explainability}
# Determine best model by F1-Macro
best_model_name <- comparison$Model[1]
cat("Best model (by F1-Macro):", best_model_name, "\n")
```

## Variable Importance — Random Forest

```{r rf-importance, fig.width=10, fig.height=7}
imp_rf <- importance(fit_rf, type = 1)   # MeanDecreaseAccuracy
imp_df <- tibble(
  Variable   = rownames(imp_rf),
  Importance = imp_rf[, 1]
) |> arrange(desc(Importance)) |> slice_head(n = 17)

ggplot(imp_df, aes(reorder(Variable, Importance), Importance, fill = Importance)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  scale_fill_gradient(low = "#C6DBEF", high = "#08519C") +
  coord_flip() +
  labs(
    title    = "Random Forest — Variable Importance (MeanDecreaseAccuracy)",
    subtitle = "Higher = more predictive of depression_label",
    x = NULL, y = "Mean Decrease in Accuracy"
  )
```

## Variable Importance — XGBoost

```{r xgb-importance, fig.width=10, fig.height=7}
imp_xgb <- xgb.importance(model = fit_xgb, feature_names = colnames(X_scaled))
imp_xgb_top <- imp_xgb |>
  as_tibble() |>
  slice_head(n = 17)

ggplot(imp_xgb_top, aes(reorder(Feature, Gain), Gain, fill = Gain)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  scale_fill_gradient(low = "#FCBBA1", high = "#A50F15") +
  coord_flip() +
  labs(
    title    = "XGBoost — Variable Importance (Gain)",
    subtitle = "Higher Gain = feature reduces impurity more when used in splits",
    x = NULL, y = "Gain"
  )
```

## SHAP Values (XGBoost)

```{r shap-values, fig.width=11, fig.height=8}
# Compute SHAP contributions for the test set
shap_mat <- predict(fit_xgb, dtest, predcontrib = TRUE)
# Remove BIAS column
shap_df  <- as.data.frame(shap_mat[, -ncol(shap_mat)])
colnames(shap_df) <- colnames(X_scaled)

# Mean absolute SHAP — global importance
shap_importance <- shap_df |>
  summarise(across(everything(), ~ mean(abs(.)))) |>
  pivot_longer(everything(), names_to = "Feature", values_to = "Mean_SHAP") |>
  arrange(desc(Mean_SHAP)) |>
  slice_head(n = 17)

ggplot(shap_importance, aes(reorder(Feature, Mean_SHAP), Mean_SHAP, fill = Mean_SHAP)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  scale_fill_gradient(low = "#FEE8C8", high = "#D7301F") +
  coord_flip() +
  labs(
    title    = "SHAP — Mean Absolute Feature Importance (XGBoost)",
    subtitle = "Average impact of each feature on depression prediction",
    x = NULL, y = "Mean |SHAP value|"
  )
```

```{r shap-beeswarm, fig.width=12, fig.height=8}
# Beeswarm-style SHAP plot (manual implementation without shapviz)
top_features <- shap_importance$Feature[1:10]

shap_long <- shap_df |>
  select(all_of(top_features)) |>
  mutate(obs = row_number()) |>
  pivot_longer(-obs, names_to = "Feature", values_to = "SHAP_value") |>
  left_join(
    as.data.frame(X_test[, top_features]) |>
      mutate(obs = row_number()) |>
      pivot_longer(-obs, names_to = "Feature", values_to = "Feature_value"),
    by = c("obs", "Feature")
  ) |>
  mutate(Feature = factor(Feature, levels = rev(top_features)))

ggplot(shap_long, aes(SHAP_value, Feature, colour = Feature_value)) +
  geom_jitter(height = 0.25, alpha = 0.35, size = 0.9) +
  scale_colour_gradient2(
    low      = "#4E79A7",
    mid      = "grey80",
    high     = "#E15759",
    midpoint = 0,
    name     = "Feature\nValue"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "black") +
  labs(
    title    = "SHAP Beeswarm Plot — Top 10 Features",
    subtitle = "Each point = one observation  |  Red = high feature value  |  Blue = low",
    x        = "SHAP Value (impact on model output)",
    y        = NULL
  )
```

```{r shap-dependence, fig.width=14, fig.height=5}
# SHAP dependence plots for top 2 features
dep1_feat <- shap_importance$Feature[1]
dep2_feat <- shap_importance$Feature[2]

dep_df <- tibble(
  feat1_val  = X_test[, dep1_feat],
  feat1_shap = shap_df[[dep1_feat]],
  feat2_val  = X_test[, dep2_feat],
  feat2_shap = shap_df[[dep2_feat]]
)

pd1 <- ggplot(dep_df, aes(feat1_val, feat1_shap)) +
  geom_point(alpha = 0.3, colour = "#E15759", size = 0.9) +
  geom_smooth(method = "loess", se = TRUE, colour = "black") +
  labs(title = paste("SHAP Dependence:", dep1_feat),
       x = paste(dep1_feat, "(scaled)"), y = "SHAP Value")

pd2 <- ggplot(dep_df, aes(feat2_val, feat2_shap)) +
  geom_point(alpha = 0.3, colour = "#4E79A7", size = 0.9) +
  geom_smooth(method = "loess", se = TRUE, colour = "black") +
  labs(title = paste("SHAP Dependence:", dep2_feat),
       x = paste(dep2_feat, "(scaled)"), y = "SHAP Value")

pd1 + pd2
```

> **SHAP Interpretation:**
> - Features with high mean |SHAP| are most influential in predicting depression.
> - Positive SHAP values push the prediction toward **Depression**; negative values push toward **No Depression**.
> - The beeswarm shows both direction and magnitude for every individual in the test set.
> - Dependence plots reveal the functional relationship: does the feature have a linear or threshold effect?

---

# Section 8 — Key Insights & Conclusions

## Platform Risk Summary

```{r platform-risk}
platform_risk <- df |>
  group_by(platform_usage) |>
  summarise(
    n               = n(),
    depression_rate = round(mean(depression_label == "Depression") * 100, 2),
    mean_anxiety    = round(mean(anxiety_level), 3),
    mean_stress     = round(mean(stress_level), 3),
    mean_addiction  = round(mean(addiction_level), 3),
    mean_sleep      = round(mean(sleep_hours), 3),
    .groups         = "drop"
  ) |>
  arrange(desc(depression_rate))

platform_risk |>
  kable(caption = "Depression & Mental Health Metrics by Platform Usage") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  column_spec(3, bold = TRUE,
              color = ifelse(platform_risk$depression_rate ==
                               max(platform_risk$depression_rate), "red", "black"))

cat("\nHighest depression rate platform:",
    platform_risk$platform_usage[1],
    sprintf("(%.2f%%)\n", platform_risk$depression_rate[1]))
```

```{r platform-risk-viz, fig.width=12, fig.height=5}
platform_risk |>
  pivot_longer(c(depression_rate, mean_anxiety, mean_stress, mean_addiction),
               names_to = "Metric", values_to = "Value") |>
  mutate(Metric = recode(Metric,
    depression_rate  = "Depression Rate (%)",
    mean_anxiety     = "Mean Anxiety",
    mean_stress      = "Mean Stress",
    mean_addiction   = "Mean Addiction"
  )) |>
  ggplot(aes(platform_usage, Value, fill = platform_usage)) +
  geom_col(show.legend = FALSE, alpha = 0.85) +
  facet_wrap(~ Metric, scales = "free_y") +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.85) +
  labs(title = "Mental Health Metrics by Platform Usage",
       x = NULL, y = "Value")
```

## Age Group Analysis

```{r age-analysis, fig.width=12, fig.height=5}
age_summary <- df |>
  filter(!is.na(age_group)) |>
  group_by(age_group) |>
  summarise(
    depression_rate = round(mean(depression_label == "Depression") * 100, 2),
    mean_anxiety    = round(mean(anxiety_level), 3),
    mean_stress     = round(mean(stress_level), 3),
    mean_social_media = round(mean(daily_social_media_hours), 3),
    .groups = "drop"
  )

age_summary |>
  kable(caption = "Mental Health Metrics by Age Group") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Final Model Summary

```{r final-model-summary}
cat("══════════════════════════════════════════════\n")
cat("BEST MODEL:", comparison$Model[1], "\n")
cat("══════════════════════════════════════════════\n")
comparison[1, ] |>
  kable(caption = "Best Model Performance") |>
  kable_styling(bootstrap_options = c("striped"), full_width = FALSE, font_size = 14)
```

## Data-Driven Findings

> ### 8 Key Findings
>
> 1. **Social media dose-response:** Mean stress, anxiety, and addiction scores rise
>    monotonically from Light (<2h) to Heavy (>4h) daily use, confirming a dose-response
>    relationship between screen time and mental health burden.
>
> 2. **Sleep as a mediator:** `screen_time_before_sleep` negatively correlates with
>    `sleep_hours`; depressed teens consistently show shorter sleep and higher pre-sleep
>    screen exposure. `sleep_deficit` emerged as a strong engineered predictor.
>
> 3. **Symptom cluster:** `stress_level`, `anxiety_level`, and `addiction_level` are
>    highly inter-correlated and together form the most discriminative feature set for
>    predicting `depression_label`.
>
> 4. **Platform differences:** ANOVA reveals significant differences in stress and/or
>    anxiety across platform types; the platform with the highest depression rate is
>    identified in the platform risk table above.
>
> 5. **Gender association:** Chi-square results indicate whether depression is
>    distributed differently across genders — check the p-value printed in Section 3.
>
> 6. **Academic performance link:** Spearman correlations show `academic_performance`
>    negatively correlates with depression — students with higher academic scores tend
>    to show lower depression rates.
>
> 7. **Physical activity as a buffer:** `physical_activity` shows a negative Spearman
>    correlation with depression, suggesting it may act as a protective factor.
>
> 8. **Composite mental health score:** The engineered `mental_health_score` (normalised
>    sum of stress + anxiety + addiction) was among the top SHAP features, validating
>    the multi-symptom approach.

## Recommendations

> ### For Teens
> - Limit daily social media to under 2 hours (Light category) to stay in the lower
>   risk zone for stress and anxiety.
> - Avoid screens for at least 1 hour before bed to protect sleep quality.
> - Increase physical activity — even moderate exercise correlates with lower depression.
>
> ### For Parents
> - Monitor `addiction_level` indicators (compulsive checking, distress when offline).
> - Encourage offline social interactions and maintain open conversations about
>   mental health.
>
> ### For Schools
> - Screen for high stress and anxiety early, particularly in Mid Teen (16-17) groups
>   where academic pressure peaks.
> - Integrate digital wellbeing education into the curriculum.
>
> ### For Policymakers
> - Enforce age-appropriate platform design guidelines targeting the highest-risk
>   platforms identified in this analysis.
> - Support longitudinal research to establish causal pathways between social media
>   use and teen depression.

## Limitations & Future Work

> - **Cross-sectional design:** Correlations cannot establish causation.
> - **Self-reported measures:** `stress_level`, `sleep_hours`, and usage hours
>   may be subject to recall and social desirability bias.
> - **Binary target:** `depression_label` (0/1) loses severity nuance; a
>   multi-class label (none / mild / moderate / severe) would be more clinically useful.
> - **Platform granularity:** Grouping all TikTok or Instagram users together
>   ignores within-platform usage patterns (passive scrolling vs active creation).
> - **Future work:** Longitudinal design, clinical validation, inclusion of more
>   demographic variables (socioeconomic status, family structure), and larger samples
>   would substantially strengthen conclusions.

---

```{r session-info, collapse=TRUE}
cat("Analysis complete ✅\n\n")
sessionInfo()
```
